Symmetric Circuits for Rank Gates

### 2018-04-06 sections/symmetric-circuits.tex
Another general edit. I have changed isomorphism-equivilent to isomorphic. I have also changed the "symmetric circuit" section in order to justify this choice of notation.

% AD - rewrote the paragraph to reduce the use of subscripts and superscripts.
Let $\tau := (R, S, \nu)$ be a many-sorted vocabulary. Let $D := \biguplus_{s
  \in S} D_{s} = \{(s,d) : d \in D_s\}$, be a disjoint union of non-empty sets,
indexed by $S$. For $R_i \in R$ let $R^D_i$ denote the full relation over the
many-sorted universe $D$, i.e.\ $R^D_i := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(R_i) = (a_1, \ldots , a_{r)})$. There is an obvious
interpretation of $\tau$ over $D$ given by assigning each sort symbol $s$ in
$\tau$ to the set $D_s$ and each relation symbol in $\tau$ to the full
relation over $D$. We call this \emph{the structure defined by $(\tau, D)$}, and
denote it by $\str{\tau, D}$.

The idea is to use the tuples in the relations of this structure as an index
set, and with the automorphisms of this structure defining a group action. Let
the \emph{index defined by the pair $(\tau, D)$} be the set $\biguplus_{R_i\in
  R} R^{D}_i := \{ (\vec{a}, R) : \vec{a} \in R^{D}, \, R \in \tau \} $. We
often use this set to index the input to a Boolean function, and as such we
denote the index defined by $(\tau, D)$ by $\ind(\tau, D)$, and call $\tau$ the
vocabulary and $D$ the universe of $\ind(\tau, D)$.

@@

% GW: move sorted permutations into the background section
For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D := \biguplus_{s \in S} D_{s}$. Let $X := \ind (\tau, D)$ and
$G \leq \sym_{X}$. We note that there is an obvious group action of $G$ on
$X$. Let $\aut (\tau, D)$ be the automorphism group of $\str{\tau, D}$. We
define a group action of $\aut (\tau, D)$ on $X$ for $\sigma \in \aut(\tau, D)$
by $\sigma \cdot \vec{a}_R := (\sigma \vec{a})_R$, for all $\vec{a}_R \in X$. We
identify $\aut(\tau, D)$ with the subgroup of $\sym_X$ that acts equivalently on
$X$.

We can extend the action of $G$ on $X$ to a right group action on the set of
functions of the form $f: X \rightarrow H$, for any set $H$, defined by $(f
\cdot \sigma)(x) := f(\sigma x)$ for all $x \in X$, $\sigma \in G$.

@@

We say that two functions $f: \ind(\tau, D) \rightarrow H$ and $g: \ind(\tau,
D') \rightarrow H$ are \emph{isomorphism-equivalent} if there exists an
isomorphism $\pi :\str{\tau, D} \rightarrow \str{\tau, D'}$ such that
$f(\vec{a}_R) = g(\pi \vec{a}_R)$ for all $\vec{a}_R \in \ind(\tau, D)$. It is
easy to see that isomorphism-equivalence is an equivalence relation.

We are often interested in the case where $X = \ind(\tau, D) = \ind(\tau, D')$.
We say that $f, g: X \rightarrow H$ are \emph{$G$-equivalent} if there exists
$\sigma \in G$ such that $f = g \sigma$. We note that in this case $f$ and $g$
are isomorphism-equivalent if, and only if, they are $\aut(\tau, D)$-equivalent.

In this paper we will often take the co-domain to be $\{0,1\}$, and so have a
function of the form $f : X \rightarrow \{0,1\}$. In this case we may think of
$f$ as defining for each relation $R^D_i$ a subset of that relation, and thus
determining a $\tau$-structure over the universe $D$. We note that two functions
$f, g: X \rightarrow \{0,1\}$ are isomorphism-equivalent if, and only if, the
two $\tau$-structures they determine are isomorphic.

We call a function of the form $F:\{0,1\}^{X} \rightarrow \{0,1\}$ a
\emph{$(\tau, D)$-structured function}, or just a \emph{structured function}
when there is no need to emphasise the particular vocabulary. We note that
$\ind(F) = \ind(\tau, D)$, justifying our choice of notation. We let \emph{the
  vocabulary of $F$} and \emph{the universe of $F$} denote $\tau$ and $D$
respectively.

@@

Structured functions give us a general framework for indexing the input sequence
to a Boolean function and associating with that function an appropriate symmetry
condition, or notion of equivalence. Let $F$ be a $(\tau, D)$-structured
function. We say $F$ is \emph{$G$-invariant} if it is constant on $G$-equivalent
input vectors, i.e.\ for all $G$-equivalent $f, g \in \{0,1\}^{X}$, $F(f) =
F(g)$. We say $F$ is \emph{isomorphism-invariant} if it is $\aut(\tau,
D)$-invariant. Note that a structured function with vocabulary $\tau$ is
isomorphism-invariant if, and only if, it determines a property of
$\tau$-structures.

@@

From the above observations, we may think of symmetric functions as
isomorphism-invariant structured functions over the single-sorted vocabulary
with a single unary relation. We can similarly think of matrix-invariant
functions as isomorphism-invariant structured functions over the appropriate
vocabulary. More formally, if the vocabulary of $F$ consists of $p$ sort symbols
$s_1, \ldots , s_p$ and only one relation symbol with type $(s_1, \ldots, s_p)$
and $F$ is isomorphism-invariant, then we say $F$ is \emph{sort-invariant}. If
$p = 2$ and $F$ is sort-invariant then we say that $F$ is
\emph{matrix-invariant}.

In keeping with convention for Boolean functions, and for introducing simplicity
without a loss of generality, we assume that the universe of any structured
function is always a disjoint union of initial segments of the natural numbers,
i.e.\ if $F$ is structured function with universe $D_{s_1} \uplus \ldots \uplus
D_{s_p}$, then for all $s \in \{s_1 , \ldots , s_p\}$ there exists $d_s \in
\nats$ such that $D_{s} = [d_s]$. We also assume that every $(\tau,
D)$-structured function has the property that every $a \in D$ appears in at
least one element of $\ind(\tau, D)$.

@@

\begin{definition}[Automorphism]\label{defn:automorphism}
  Let $C = \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B},\tau)$-circuit of order $n$ computing a $q$-ary query, and where
  $\mathbb{B}$ is a basis of isomorphism-invariant structured functions. Let
  $\sigma \in \sym_n$ and $\pi: G \rightarrow G$ be a bijection such that
  \begin{myitemize}
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, $\Sigma (g) = \Sigma (\pi g)$,
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$, and
  \item For each pair of gates $g, h \in G$ $W(h,g)$ if and only if $W(\pi h,
    \pi g)$ and for each internal gate $g$ we have that $L(\pi g)$ and $ \pi
    \cdot L(g)$ are isomorphism-equivalent.
  \end{myitemize}
  We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
  \emph{extends to an automorphism} $\pi$. The group of automorphisms of $C$ is
  called $\aut (C)$.
\end{definition}
We can equally define an isomorphism between a pair of circuits $C =
\langle G, \Omega, \Sigma, \Lambda, L\rangle$ and  $C' = \langle G',
\Omega', \Sigma', \Lambda', L'\rangle$ as a bijection $\pi: G
\rightarrow G'$ satisfying conditions as above.  We do not usually
need to consider distinct, isomorphic circuits and for this reason we
only formally define automorphsims.

@@

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. We recursively define the equivalence relation
  \emph{syntactic-equivalence}, which we denote using the symbol `$\equiv$', on
  $G$ as follows. Suppose $g$ and $h$ are gates in $C$ such that $\Sigma (g) =
  \Sigma(h)$ and either both $g$ and $h$ are output gates and $\Omega^{-1}(g) =
  \Omega^{-1}(h)$ or neither are output gates. Suppose $g$ and $h$ are input
  gates, then $g \equiv h$ if, and only if, both $g$ and $h$ are constant gates
  or both are relational gates and $\Lambda(g) = \Lambda (h)$. Suppose $g$ and
  $h$ are internal gates and suppose we have defined the syntactic-equivalence
  relation for all gates of depth less than the depth of either $g$ or $h$. Then
  $g \equiv h$ if, and only if, $L(g) /_\equiv$ and $L(h) /_\equiv$ are
  isomorphism-equivalent.
\end{definition}

% \begin{drem}
%   Should I prove these are actually equivalence relations? I use that fact, but
%   it's very easy to prove.
% \end{drem}

@@

  Suppose $g$ is an internal gate and suppose that for every gate $h$ of depth
  less than $g$, $\pi (h)$ is syntactically-equivalent to $\pi'(h)$. We have
  that there exists $\lambda, \lambda' \in \aut(g)$ such that $\pi
  L(g)(\vec{a}_R) = L(\pi g) (\lambda \vec{a}_R)$ and $\pi' L(g)(\vec{a}_R) =
  L(\pi' g)(\lambda' \vec{a}_R)$, for all $\vec{a}_R \in ind(g)$. Then, from the
  inductive hypothesis, we have that $\pi L(g)(\vec{a}_R) \equiv \pi
  'L(g)(\vec{a}_R)$ and so $L(\pi g)(\lambda \vec{a}_R) \equiv L(\pi' g)
  (\lambda' \vec{a}_R)$, for all $\vec{a}_R \in \ind(g)$. Thus we have that
  $L(\pi g) (\vec{a}_R) = L (\pi g) (\lambda \lambda^{-1}\vec{a}_R) \equiv
  L(\pi' g) (\lambda' \lambda^{-1}\vec{a}_R)$, for all $\vec{a}_R \in \ind(g)$,
  and so $L(\pi g)$ is isomorphism equivalent to $L(\pi'g)$ up to
  syntactic-equivalence. We thus have that $\pi (g) $ and $\pi' (g)$ are
  syntactically-equivalent, and the result follows.

@@

We have from Proposition~\ref{prop:unique-children-unique-extensions} that
circuits with unique gates have unique extensions, but, in order to prove many
of the results we will need later, we not only need that each automorphism is
uniquely defined by the permutation it extends, but also that the isomorphisms
that witness the isomorphism-equivalences, and so witness the validity of the
automorphism, are also uniquely defined ( or, at the very least, well
constrained) by the permutation $\sigma$. Without a sufficient constraint on the
choice of isomorphism, testing if two gates have isomorphism-equivalent
labellings (a requirement for computing the automorphism extending a
permutation) requires solving some version of the isomorphism problem for
$\tau$-structures, where $\tau$ is the vocabulary of the gates in question.
Indeed, we prove in Section~\ref{sec:transparent}, that for the class of
circuits with unique children deciding if there is an isomorphism-equivalence
between two gates is as hard as the graph-isomorphism problem. It follows that a
great many important circuit properties (e.g.\ the orbit of a gate, symmetry of
a circuit, etc.) for circuits with unique gates are at the very least not
obviously polynomial-time decidable. As such, circuits with unique children,
although satisfying the requirement of having unique extensions, are
insufficient for our purposes.

@@

In order to ensure a canonical choice of isomorphism we instead consider
circuits with unique labels. It is easy to see that if $C$ has unique labels
then $L(g)$ and $L(g')$ are isomorphism-equivalent if, and only if, $\lambda :=
L(g')^{-1}L(g)$ is an isomorphism. In this case, rather than having to decide
the isomorphism problem, we can decide isomorphism-equivalences by checking if a
given candidate function is a valid isomorphism. We prove in
Section~\ref{sec:transparent} that all of the circuit properties of interest in
this paper are polynomial-time decidable for circuits with unique labels.
Moreover, we show in Lemma~\ref{lem:transparent-unique} that a transparent
circuit can be transformed into an equivalent circuit with unique labels in time
polynomial in the size of the circuit. Since we are working with $P$-uniform
families of transparent circuits, this fact allows us to assume, without a loss
of generality, that transparent circuits have unique labels and so , from
\ref{prop:unique-children-unique-extensions}, unique extensions).

@@

Before we state the main result of this paper we need to discuss the Boolean
bases of interest in this paper and introduce some useful terminology. In most
other contexts a basis is taken to be a finite sequence of symmetric Boolean
functions, and a binary encoding of a circuit may is given by assigning to each
element of the basis a unique symbol, and then encoding the circuit as a
labelled DAG. Moreover, since there are only a finite number of basis elements,
it follows that if each function in the basis can be computed in polynomial time
then the mapping that takes a symbol $s$ and a string $\vec{x}$ to the value of
the function denoted by $s$ for input $\vec{x}$ can also be decided in
polynomial-time. We call this function the \emph{gate evaluation function}. It
is easy to see that the gate evaluation function is polynomial-time computable
if, and only if, the a given circuit can be evaluated in polynomial-time. It is
here our that our definition of a basis introduces a complication. In order to
allow for the inclusion of structured (possibly non-symmetric) functions in the
basis, we have instead defined a basis to be a (possibly infinite) set of finite
Boolean functions. In this case the gate evaluation function may not even be
decidable, let alone polynomial-time decidable.

@@

\begin{thm}[Main Theorem]
  A graph property is decidable by a $\P$-uniform family of transparent
  symmetric rank-circuits if, and only if, it is definable by an $\FPR$
  sentence.
\end{thm}

>>>>

Let $\tau := (R, S, \nu)$ be a many-sorted vocabulary. Let $D := \biguplus_{s
  \in S} D_{s} = \{(s,d) : d \in D_s\}$, be a disjoint union of non-empty sets,
indexed by $S$. For $R_i \in R$ let $R^D_i$ denote the full relation over the
many-sorted universe $D$, i.e.\ $R^D_i := D_{a_1} \times \ldots \times
D_{a_{r}}$, where $\nu(R_i) = (a_1, \ldots , a_{r)})$. There is an obvious
interpretation of $\tau$ over $D$ given by assigning each sort symbol $s$ in
$\tau$ to the set $D_s$ and each relation symbol in $\tau$ to the full relation
over $D$. We call this \emph{the structure defined by $(\tau, D)$}, and denote
it by $\str{\tau, D}$.

The idea is to use the tuples in the relations of this structure as an index
set, and with the automorphisms of this structure defining a group action. Let
the \emph{index defined by the pair $(\tau, D)$} be the set $\biguplus_{R_i\in
  R} R^{D}_i := \{ (\vec{a}, R) : \vec{a} \in R^{D}, \, R \in \tau \} $. We
often use this set to index the input to a Boolean function, and as such we
denote the index defined by $(\tau, D)$ by $\ind(\tau, D)$. We call $\tau$ the
vocabulary and $D$ the universe of $\ind(\tau, D)$.

@@

For the remainder of this subsection we fix a vocabulary $\tau$ and a
many-sorted set $D := \biguplus_{s \in S} D_{s}$, and let $X := \ind (\tau, D)$
and $G \leq \sym_{X}$. There is an obvious group action of $G$ on $X$. Let $\aut
(\tau, D)$ be the automorphism group of $\str{\tau, D}$. We define a group
action of $\aut (\tau, D)$ on $X$ for $\sigma \in \aut(\tau, D)$ by $\sigma
\cdot \vec{a}_R := (\sigma \vec{a})_R$ for all $\vec{a}_R \in X$. We identify
$\aut(\tau, D)$ with the subgroup of $\sym_X$ that acts equivalently on $X$.

Let $H$ be a set. We think of a function $f : X \rightarrow H$ as defining a
labelling of $\str{\tau,D}$ by $H$, and we identify $f$ with this labelled
instance of $\str{\tau, D}$. Let $f: \ind(\tau, D) \ra H$ and $g : \ind (\tau,
D') \ra H$. We say that $f$ and $g$ are \emph{isomorphic} if there is an
isomorphism $\pi : \str{\tau, D} \ra \str{\tau, D'}$ such that $f(\vec{a}_R) =
g(\pi(\vec{a}_R))$ for all $\vec{a}_R \in \ind(\tau, D)$. In other words, $f$
and $g$ are isomorphic if, and only if, they are isomorphic as (labelled)
structures. We will often take the co-domain $H$ to be $\{0,1\}$. In this case
$f$ and $g$ can be thought of as each defining a sub-structure of
$\str{\tau,D}$. It is easy to see that $f$ and $g$ are isomorphic
if, and only if, the $\tau$ structures they define are isomorphic.

@@

% We say that two functions $f: \ind(\tau, D) \rightarrow H$ and $g: \ind(\tau,
% D') \rightarrow H$ are \emph{isomorphism-equivalent} if there exists an
% isomorphism $\pi :\str{\tau, D} \rightarrow \str{\tau, D'}$ such that
% $f(\vec{a}_R) = g(\pi \vec{a}_R)$ for all $\vec{a}_R \in \ind(\tau, D)$. It is
% easy to see that isomorphism-equivalence is an equivalence relation.

% We are often interested in the case where $X = \ind(\tau, D) = \ind(\tau,
% D')$. We say that $f, g: X \rightarrow H$ are \emph{$G$-equivalent} if there
% exists $\sigma \in G$ such that $f = g \sigma$. We note that in this case $f$
% and $g$ are isomorphism-equivalent if, and only if, they are $\aut(\tau,
% D)$-equivalent.

% In this paper we will often take the co-domain to be $\{0,1\}$, and so have a
% function of the form $f : X \rightarrow \{0,1\}$. In this case we may think of
% $f$ as defining for each relation $R^D_i$ a subset of that relation, and thus
% determining a $\tau$-structure over the universe $D$. We note that two
% functions $f, g: X \rightarrow \{0,1\}$ are isomorphism-equivalent if, and
% only if, the two $\tau$-structures they determine are isomorphic.

We say that $f, g: X \rightarrow H$ are \emph{$G$-equivalent} if there exists
$\sigma \in G$ such that $f = g \sigma$. We note that $f$ and $g$ are isomorphic
if, and only if, they are $\aut(\tau, D)$-equivalent.

We call a function of the form $F:\{0,1\}^{X} \rightarrow \{0,1\}$ a
\emph{$(\tau, D)$-structured function}, or just a \emph{structured function}. We note that
$\ind(F) = \ind(\tau, D)$, justifying our choice of notation. We let \emph{the
  vocabulary of $F$} and \emph{the universe of $F$} be $\tau$ and $D$,
respectively.

@@

It is important to note that structured functions not only provide us with way
of talking about Boolean functions that take as input (encodings of) structures,
but their definition also suggests a natural notion of symmetry. Let $F$ be a
$(\tau, D)$-structured function. We say $F$ is \emph{$G$-invariant} if for all
$G$-equivalent $f, g : X \ra \{0,1\}$, $F(f) = F(g)$. We say $F$ is
\emph{isomorphism-invariant} if $F$ is $\aut(\tau, D)$-invariant. Note that a
structured function with vocabulary $\tau$ is isomorphism-invariant if, and only
if, it decides a property of $\tau$-structures.

@@

We may think of symmetric functions as isomorphism-invariant structured
functions over the single-sorted vocabulary with a single unary relation. We can
similarly think of matrix-invariant functions as isomorphism-invariant
structured functions over an appropriate vocabulary. More formally, if the
vocabulary of $F$ consists of $p$ sort symbols $s_1, \ldots , s_p$ and only one
relation symbol with type $(s_1, \ldots, s_p)$ and $F$ is isomorphism-invariant,
then we say $F$ is \emph{sort-invariant}. If $p = 2$ and $F$ is sort-invariant
then we say that $F$ is \emph{matrix-invariant}.

In keeping with the conventions for Boolean functions we usually assume, without
a loss of generality, that the universe of a structured function is a disjoint
union of initial segments of the natural numbers, i.e.\ if $F$ is structured
function with universe $D_{s_1} \uplus \ldots \uplus D_{s_p}$ then for all $s
\in \{s_1 , \ldots , s_p\}$ there exists $d_s \in \nats$ such that $D_{s} =
[d_s]$. We also assume that every $(\tau, D)$-structured function has the
property that every $a \in D$ appears in at least one element of $\ind(\tau,
D)$.

@@

\begin{definition}[Automorphism]\label{defn:automorphism}
  Let $C = \langle G, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B},\tau)$-circuit of order $n$ computing a $q$-ary query, and where
  $\mathbb{B}$ is a basis of isomorphism-invariant structured functions. Let
  $\sigma \in \sym_n$ and $\pi: G \rightarrow G$ be a bijection such that
  \begin{myitemize}
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, $\Sigma (g) = \Sigma (\pi g)$,
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$, and
  \item For each pair of gates $g, h \in G$ $W(h,g)$ if and only if $W(\pi h,
    \pi g)$ and for each internal gate $g$ we have that $L(\pi g)$ and $ \pi
    \cdot L(g)$ are isomorphic.
  \end{myitemize}
  We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
  \emph{extends to an automorphism} $\pi$. The group of automorphisms of $C$ is
  called $\aut (C)$.
\end{definition}
We can equally define an isomorphism between a pair of circuits $C = \langle G,
\Omega, \Sigma, \Lambda, L\rangle$ and $C' = \langle G', \Omega', \Sigma',
\Lambda', L'\rangle$ as a bijection $\pi: G \rightarrow G'$ satisfying
conditions as above. We do not usually need to consider distinct, isomorphic
circuits and for this reason we only formally define automorphsims.

@@

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ be a $(\mathbb{B},
  \rho)$-circuit of order $n$. We recursively define the equivalence relation
  \emph{syntactic-equivalence}, which we denote using the symbol `$\equiv$', on
  $G$ as follows. Suppose $g$ and $h$ are gates in $C$ such that $\Sigma (g) =
  \Sigma(h)$ and either both $g$ and $h$ are output gates and $\Omega^{-1}(g) =
  \Omega^{-1}(h)$ or neither are output gates. Suppose $g$ and $h$ are input
  gates, then $g \equiv h$ if, and only if, both $g$ and $h$ are constant gates
  or both are relational gates and $\Lambda(g) = \Lambda (h)$. Suppose $g$ and
  $h$ are internal gates and suppose we have defined the syntactic-equivalence
  relation for all gates of depth less than the depth of either $g$ or $h$. Then
  $g \equiv h$ if, and only if, $L(g) /_\equiv$ and $L(h) /_\equiv$ are
  isomorphic.
\end{definition}

% \begin{drem}
%   Should I prove these are actually equivalence relations? I use that fact,
%   but it's very easy to prove.
% \end{drem}

@@

  Suppose $g$ is an internal gate and suppose that for every gate $h$ of depth
  less than $g$, $\pi (h)$ is syntactically-equivalent to $\pi'(h)$. We have
  that there exists $\lambda, \lambda' \in \aut(g)$ such that $\pi
  L(g)(\vec{a}_R) = L(\pi g) (\lambda \vec{a}_R)$ and $\pi' L(g)(\vec{a}_R) =
  L(\pi' g)(\lambda' \vec{a}_R)$, for all $\vec{a}_R \in ind(g)$. Then, from the
  inductive hypothesis, we have that $\pi L(g)(\vec{a}_R) \equiv \pi
  'L(g)(\vec{a}_R)$ and so $L(\pi g)(\lambda \vec{a}_R) \equiv L(\pi' g)
  (\lambda' \vec{a}_R)$, for all $\vec{a}_R \in \ind(g)$. Thus we have that
  $L(\pi g) (\vec{a}_R) = L (\pi g) (\lambda \lambda^{-1}\vec{a}_R) \equiv
  L(\pi' g) (\lambda' \lambda^{-1}\vec{a}_R)$, for all $\vec{a}_R \in \ind(g)$,
  and so $L(\pi g)/_\equiv$ is isomorphic to $L(\pi'g) /_\equiv$. We thus have
  that $\pi (g) $ and $\pi' (g)$ are syntactically-equivalent, and the result
  follows.

@@

We have from Proposition~\ref{prop:unique-children-unique-extensions} that
circuits with unique gates have unique extensions. However, in order to prove
many of the results we will need later we not only need that each automorphism
is uniquely determined by the permutation it extends, but also that the
isomorphisms between the labellings of each gate that witness 
this automorphism are also uniquely determined by the permutation -- or, at the
very least, well constrained. Without this constraint, computing many properties
of the circuit would involve solving some version of the isomorphism problem for
$\tau$-structures, where $\tau$ is the vocabulary of the gates in question.
Indeed, we prove in Section~\ref{sec:transparent}, that for the class of
circuits with unique children deciding if a a given a given function is in fact
a valid circuit automorphism is at least as hard as hard as the
graph-isomorphism problem. Moreover, we go on to show that a great many
important circuit 
properties (e.g.\ the orbit of a gate, symmetry of a circuit, etc.) are also at
least as hard to decide for circuits with unique gates as the graph isomorphism
-- and so these properties 
are, at the very least, not obviously polynomial-time decidable. As such,
circuits with unique children, although satisfying the requirement of having
unique extensions, are insufficient for our purposes.

@@

In order to ensure a canonical choice of isomorphism we instead consider
circuits with unique labels. It is easy to see that if $C$ has unique labels
then $L(g)$ and $L(g')$ are isomorphic if, and only if, $\lambda :=
L(g')^{-1}L(g)$ is an automorphism of $\str{g'}$ . We see that in this case the problem of
deciding if $L(g)$ and $L(g')$ are isomorphic reduces to the problem of deciding if a given
candidate function is a valid automorphism, a much easier problem. Indeed, we
prove in Section~\ref{sec:transparent} that all of the circuit properties of
interest in this paper are polynomial-time decidable for circuits with unique
labels. Moreover, we show in Lemma~\ref{lem:transparent-unique} that a
transparent circuit can be transformed into an equivalent circuit with unique
labels in time polynomial in the size of the circuit. Since we are working with
$\PT$-uniform families of transparent circuits, this fact allows us to assume,
without a loss of generality, that transparent circuits have unique labels and
so, from Proposition~\ref{prop:unique-children-unique-extensions}, unique extensions.

@@

Before we state the main result of this paper we need to discuss some of the Boolean
bases that will be of interest to us and introduce some useful terminology. In most
other contexts a basis is taken to be a finite sequence of symmetric Boolean
functions, and a binary encoding of a circuit may is given by assigning to each
element of the basis a unique symbol, and then encoding the circuit as a
labelled DAG. Moreover, since there are only a finite number of basis elements,
it follows that if each function in the basis can be computed in polynomial time
then the mapping that takes a symbol $s$ and a string $\vec{x}$ to the value of
the function denoted by $s$ for input $\vec{x}$ can also be decided in
polynomial-time. We call this function the \emph{gate evaluation function}. It
is easy to see that the gate evaluation function is polynomial-time computable
if, and only if, the a given circuit can be evaluated in polynomial-time. It is
here our that our definition of a basis introduces a complication. In order to
allow for the inclusion of structured (possibly non-symmetric) functions in the
basis, we have instead defined a basis to be a (possibly infinite) set of finite
Boolean functions. In this case the gate evaluation function may not even be
decidable, let alone polynomial-time decidable.

@@

\begin{thm}[Main Theorem]
  A graph property is decidable by a $\PT$-uniform family of transparent
  symmetric rank-circuits if, and only if, it is definable by an $\FPR$
  sentence.
\end{thm}
